---
title: "NickBlackfordWeek10"
author: "Nick Blackford"
date: "`r Sys.Date()`"
output: pdf_document
---
## Thoracic Surgery Dataset 
```{r}
library(foreign)

mydata <- read.arff("/Users/nickblackford/Downloads/ThoraricSurgery.arff")

head(data)
```
```{r}
library(MASS)
# Multivariate logistic regression with stepwise selection
full_model <- glm(Risk1Yr ~ PRE7 + PRE8 + PRE9 +PRE10 + PRE11 + PRE17 +PRE19 +PRE25 + PRE30 + PRE32, family = binomial, data = mydata)
step_model <- stepAIC(full_model, direction = "both")
summary(step_model)
```
```{r}

#best model per stepwise selection 
model1 <- glm(formula = Risk1Yr ~ PRE9 + PRE11 + PRE17 + PRE30, family = binomial, 
    data = mydata)
summary(model1)
```
PRE9T (1.1754): This variable has the largest coefficient, indicating that it has the strongest positive effect on the survival rate. It is statistically significant at the 0.01 level

PRE17T (0.9940): This variable has the second-largest coefficient, suggesting a strong positive effect on the survival rate. It is statistically significant at the 0.05 level 
PRE30T (0.8406): This variable has a positive effect on the survival rate, but it is only marginally significant at the 0.05 level.

PRE11T (0.5232): This variable has the smallest coefficient among the significant predictors, indicating a weaker positive effect on the survival rate. It is not statistically significant at the 0.05 level.

```{r}
#  Predict probabilities
predicted_probabilities <- predict(model1, type = "response")

# Convert probabilities to binary predictions
threshold <- 0.5
predicted_outcomes <- ifelse(predicted_probabilities > threshold, 1, 0)

# Calculate accuracy
actual_outcomes <- mydata$Risk1Yr
actual_outcomes_numeric <- as.numeric(actual_outcomes) - 1
correct_predictions <- sum(predicted_outcomes == actual_outcomes_numeric)
total_predictions <- length(predicted_outcomes)
accuracy <- correct_predictions / total_predictions

# Print the accuracy
print(accuracy)
```
Accuracy for my model was 85% with a threshold of 0.5 to shift my probabilities into binary results. 

## Binary Classifier Data 
```{r, results='hide', echo=FALSE}
library(dplyr)
library(tidyverse)
classifier_data <- read_csv('/Users/nickblackford/Downloads/binary-classifier-data.csv')
```
```{r}

#logistic regression model
log_model <- glm(formula = label ~ x + y, data = classifier_data)
summary(log_model)
```
```{r}
#  Predict probabilities
predicted_probabilities_2 <- predict(log_model, type = "response")

# Convert probabilities to binary predictions
threshold_2 <- 0.5
predicted_outcomes_2 <- ifelse(predicted_probabilities_2 > threshold_2, 1, 0)

# Calculate accuracy
actual_outcomes_2<- classifier_data$label
correct_predictions_2 <- sum(predicted_outcomes_2 == actual_outcomes_2)
total_predictions_2 <- length(predicted_outcomes_2)
accuracy_2 <- correct_predictions_2 / total_predictions_2

# Print the accuracy
print(accuracy_2)
```
This model had an accuracy of 58% when comparing predicted to actual values. 


